---
title: "An Analysis on the status of Visa Cases"
output: pdf_document
---

Created By: Nirvan Silswal 
NetID: ns318
Email: nirvan.silswal@duke.edu

##Load Library Packages

```{r}
library(tidyverse)
library(readxl)
VisaData <- read_excel("DIIG F20 Data Challenge #2.xlsx")

```

In this dataset we have data on 167,278 different visa applications each with 16 different attributes associated with the application. 

During this analysis I want to answer two major questions:

1. What variables makes an application more likely to get approved and what variables make an application less likely to get approved. 

2. How do Job wages compare across locations?


Lets look at question 1 first:



```{r - Question1}

```

Now lets look at how we can answer question 2:

To analyze wages lets first construct a plot of all the different jobs in the dataset

```{r - Question 2}
ggplot(data = VisaData, mapping = aes(x = JOB_TITLE_SUBGROUP)) + scale_x_discrete(guide=guide_axis(n.dodge=2)) +
  geom_bar()
```

Now lets look at the locations where these visa applicants live:

```{r}
ggplot(data = VisaData, mapping = aes(x = WORK_STATE_ABBREVIATION)) + scale_x_discrete(guide=guide_axis(n.dodge=10)) +
  geom_bar()

```


It's clear to see that the overwhelming majority of Visa-Applicants in this dataset are residing in California.
This is important to note as California is a hub for software development jobs. Lets take a look at how many people who applied for a Visa in California also have a software related job.

```{r}
ggplot(data = VisaData, mapping = aes(x = (WORK_STATE = "California"), fill = JOB_TITLE_SUBGROUP)) + scale_x_discrete(guide=guide_axis(n.dodge=10)) +
  geom_bar()
```

An overwhelming majority of the applicants from California are working some sort of software job. This is important to note as these software related jobs typically pay much more than say a teacher. 

To further analyze this lets take a look at a Box plots of wages in each state

```{r}
ggplot(data = VisaData, mapping = aes(y = PAID_WAGE_PER_YEAR, x = WORK_STATE_ABBREVIATION)) +
  scale_x_discrete(guide=guide_axis(n.dodge=10)) +
  geom_boxplot()
```















