---
title: "Analyzing Visa Applicant Demographics"
output: pdf_document
---

Created By: Nirvan Silswal 
NetID: ns318
Email: nirvan.silswal@duke.edu

##Load Library Packages

```{r}
library(tidyverse)
library(readxl)
VisaData <- read_excel("DIIG F20 Data Challenge #2.xlsx")

```



In this dataset we have data on 167,278 different visa applications each with 16 different attributes associated with the application. 

During this analysis I want to answer two major questions:

1. What variables makes an application more likely to get approved and what variables make an application less likely to get approved. 

2. How do Job wages compare across locations?


Lets look at question 1 first:

To start off, we should look at where are applicants who get certified apply from, and where applicants who are denied apply from.

```{r - Question1}
 
  ggplot(data = VisaData, mapping = 
      aes(x = (WORK_STATE_ABBREVIATION), fill = CASE_STATUS)) +   
      geom_bar(position = "fill") + scale_x_discrete(guide=guide_axis(n.dodge=10))
  labs(y = "proportion")

```
Certification Rate = the percentage of applicants who were Certified by the Visa office.

While most states hover around and 80% Certification rate, it is interesting to note that the US territory of the Northern Marina Islands (MP) has a Certification rate of less than 50%. This is likely due to the fact that MP is a US territory and not a state - inticing Visa offices to approve less applicants from there.

For the most part, for those applying from a US state, there is no significant difference between Visa certifiaciton rate between states.

It might be more beneficial to analyze certification rates based on the job an applicant has. Lets take a look at that now:

```{r}

ggplot(data = VisaData, mapping = 
      aes(x = (JOB_TITLE_SUBGROUP), fill = CASE_STATUS)) +   
      geom_bar(position = "fill") + scale_x_discrete(guide=guide_axis(n.dodge=10))
  labs(y = "proportion")

```
Based on the data, there doesn't appear to be a signifcant difference regarding the occupation an applicant holds and their probability of being approved. All the jobs here seem to fluctuate between a 75% - 92% Certificaiton rate. It is worth noting that almost every job had a certificaiton rate of about 92% except for Asistant professors and attorney's - those were closer to the 75% Certification rate.

Finally, we can take a look at how the Visa Class applied for influences the certification rate for an applicant.

```{r}

ggplot(data = VisaData, mapping = 
      aes(x = (VISA_CLASS), fill = CASE_STATUS)) +   
      geom_bar(position = "fill") + scale_x_discrete(guide=guide_axis(n.dodge=10))
  labs(y = "proportion")

```

Here we can clearly see that applicants who applied for a Greencard were certified at a rate of about 60% - much lower than the other Visa Classes. Furthermore H-1B1 applicants from Singapore were approved at the highest Certification Rate - almost 90%.

In summary: 

Now lets look at how we can answer question 2:

To analyze wages lets first construct a plot of all the different jobs in the dataset

```{r - Question 2}
ggplot(data = VisaData, mapping = aes(x = JOB_TITLE_SUBGROUP)) + scale_x_discrete(guide=guide_axis(n.dodge=2)) +
  geom_bar()
```

Now lets look at the locations where these visa applicants live:

```{r}
ggplot(data = VisaData, mapping = aes(x = WORK_STATE_ABBREVIATION)) + scale_x_discrete(guide=guide_axis(n.dodge=10)) +
  geom_bar()

```


It's clear to see that the overwhelming majority of Visa-Applicants in this dataset are residing in California.
This is important to note as California is a hub for software development jobs. Lets take a look at how many people who applied for a Visa in California also have a software related job.

```{r}
ggplot(data = VisaData, mapping = aes(x = (WORK_STATE = "California"), fill = JOB_TITLE_SUBGROUP)) + scale_x_discrete(guide=guide_axis(n.dodge=10)) +
  geom_bar()
```

An overwhelming majority of the applicants from California are working some sort of software job. This is important to note as these software related jobs typically pay much more than say a teacher. 





















To further analyze this we should look at average wages in each state:

```{r, fig.height=5, fig.width=5}

WageMean <- aggregate( PAID_WAGE_PER_YEAR ~ WORK_STATE, VisaData, mean )
WageMean <- WageMean[order(WageMean$PAID_WAGE_PER_YEAR,decreasing=T),]
WageMean

```















